---
title: <span style="color:white">TuneR</span>
author: <span style="color:white"> JHCV </span>
date: <span style="color:white"> 2023-04-25 </span>
output: html_document
---

<body style="background-color:#02233F">
<font style="color:white">

```{css, echo=FALSE, message = FALSE, warning = FALSE}
.jhcv {
  background-color: rgb(159, 199, 219);
  border: 3px solid white;
  font-weight: bold;
}
```

```{r setup, include=TRUE, class.source="jhcv", warning = FALSE, message = FALSE}
timeStarted <- Sys.time()
knitr::opts_chunk$set(echo = TRUE)
##### Required packages #####
require(tidyverse)
require(lubridate)
require(plotly)
require(spotifyr)
require(ggwordcloud)
require(gganimate)
require(gifski)
require(purrr)
require(kableExtra)
#####
##### Plot Appearance Theme #####
plotartitmyTheme <- theme(legend.position = "none",
                 plot.background = element_rect(fill = "#02233F"),
                 panel.background = element_rect(fill = "#02233F"),
                 panel.grid = element_line(color = "#274066"),
                 axis.text = element_text(color = "white"),
                 axis.title = element_text(color = "white",
                                           hjust = .5),
                 plot.title = element_text(color = "white",
                                           hjust = .5),
                 plot.subtitle = element_text(color = "white",
                                              hjust = .5),
                 strip.background = element_rect(fill = "#02233F"),
                 strip.text.x = element_text(color = "white"),
                 plot.caption = element_text(color = "white",
                                             size = 5))
#####
##### Legend appearance theme #####
myLegend <- theme(legend.position = "right",
                  legend.background = element_rect(fill = "#02233F"),
                  legend.text = element_text(color = "white"),
                  legend.title = element_text(color = "white",
                                              hjust = .5))
#####

##### Spotify API Authorization #####

Sys.setenv("SPOTIFY_CLIENT_ID" = "xxx")
Sys.setenv("SPOTIFY_CLIENT_SECRET" = "xxx")
Sys.setenv("SPOTIFY_REDIRECT_URI" = "http://localhost:1410/")

accessToken <- get_spotify_access_token(client_id = Sys.getenv("SPOTIFY_CLIENT_ID"), 
                                        client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET"))

##### 

```

# What have you been listening to lately?

```{r, class.source="jhcv", message = FALSE}

# Create function for iterating over time frames

plot_genre_wordcloud <- function(timeframe = "short") {
  
  data <- get_my_top_artists_or_tracks("artists",
                               time_range = paste(timeframe,
                                                  "_term",
                                                  sep = ""),
                               limit = 50) |>
  
  select(name, genres, popularity) |> 
  
  unnest(genres) |>
  
  select(genres) |>
  
  group_by(genres) |>
  
  summarize("n" = n())
  
  ggplot(data,
         aes(label = genres,
             size = n,
             color = factor(genres))) +
  
  geom_text_wordcloud(shape = "diamond") +
  
  myTheme
  
}

plot_artist_wordcloud <- function(timeframe = "long") {
  
  data <- get_my_top_artists_or_tracks("artists",
                                       time_range = paste(timeframe,
                                                          "_term",
                                                          sep = ""),
                                       limit = 50,
                                       offset = 0) |>
    
    select(name, popularity, genres) 
  
  data$sf <- 0
  
  for (i in 1:nrow(data)) {
    
    data$sf[i] <- (51-i)^1.2
    
  }
  
  ggplot(data,
         aes(label = name,
             size = sf,
             color = factor(name))) +
    
    geom_text_wordcloud(shape = "diamond") +
    
    myTheme
  
}

```

## Short Term

Calculated using data from the last 4 weeks. 

```{r, class.source="jhcv", message = FALSE}

mtg <- plot_genre_wordcloud("short_term")

mtg

```

## Medium Term

Calculated using data from the last 6 months.

```{r, class.source="jhcv", message = FALSE}

mtgwc <- plot_genre_wordcloud("medium_term")

mtgwc

```

## Long Term

Calculated using several years of data including all new data as it becomes available.

```{r, class.source="jhcv", message = FALSE}

ltgwc <- plot_genre_wordcloud("long_term")

ltgwc

```

# Pull All Saved Tracks

```{r, class.source="jhcv", message = FALSE}

# Retrieve all saved tracks
my_saved_tracks <- NULL
offset <- 0
saved_track_count <- 0
repeat {
  track_data <- get_my_saved_tracks(limit = 50, 
                                    offset = offset) |>
    select(added_at,
           track.artists,
           track.duration_ms,
           track.explicit,
           track.id,
           track.name,
           track.popularity,
           track.album.id,
           track.album.images,
           track.album.name,
           track.album.release_date)
  my_saved_tracks <- rbind(my_saved_tracks, track_data)
  saved_track_count <- saved_track_count + nrow(track_data)
  message(paste0("Saved Tracks: ", 
               saved_track_count, 
               "/x"))
  if (nrow(track_data) < 50) {
    break
  }
  offset <- offset + 50
}

my_saved_tracks$added_at <- as_datetime(my_saved_tracks$added_at,
                                        tz = Sys.timezone())

my_saved_tracks$track.album.release_date <- as.Date(my_saved_tracks$track.album.release_date)

for (i in 1:nrow(my_saved_tracks)) {
  
  my_saved_tracks$track.artists[[i]] <- my_saved_tracks$track.artists[[i]]$name[1]
  
  if (my_saved_tracks$track.popularity[i] == 0) {
    
    my_saved_tracks$track.popularity[i] <- 0.1
    
  }
  
}

my_saved_tracks$track.artists <- unlist(my_saved_tracks$track.artists)

# - Fixing Outliers - Buju Banton realeased "Hills and Valleys" in 1997, not 1900 lol

my_saved_tracks <- my_saved_tracks |> 
  mutate(track.album.release_date = case_when(
    track.artists == "Buju Banton" & track.name == "Hills And Valleys" ~ as.Date("1997-11-18"),
    TRUE ~ track.album.release_date  # Keeps existing dates unchanged
  ))

```

# Plot Release Date Histogram

```{r, class.source="jhcv", warning = FALSE}

mstp <- ggplot(my_saved_tracks,
               aes(x = as.Date(track.album.release_date),
                   color = factor(track.artists),
                   weight = ..density..)) +
  
  geom_histogram(fill = "white",
                 bins = 100) +
  
  labs(x = "Release Date",
       y = "Density",
       title = "Release Date Histogram") +
  
  myTheme

ggplotly(mstp)

```

# When did you get to the party?

```{r, class.source="jhcv"}

mst <- my_saved_tracks |> 
  
  select(added_at,
         track.artists,
         track.name,
         track.album.release_date,
         track.popularity) |>
  
  drop_na() |>
  
  arrange(added_at)

msttsrda <- ggplot(mst,
                   aes(x = as_datetime(added_at,
                                tz = Sys.timezone()),
                       y = as.Date(track.album.release_date),
                       color = factor(track.name),
                       fill = factor(track.artists),
                       alpha = track.popularity/100)) +
  
  geom_point() +
  
  labs(x = "Date Saved",
       y = "Release Date",
       title = "Discovery Scatterplot") +
  
  scale_alpha_identity() +
  
  myTheme +
  
  transition_reveal(as_datetime(added_at,
                                tz = Sys.timezone())) #+
  #ease_aes("elastic-in-out") +
  #enter_grow()

animate(msttsrda, renderer = gifski_renderer())

```

```{r, class.source="jhcv"}

msttsrdp <- ggplot(my_saved_tracks,
                   aes(x = added_at,
                       y = as.Date(track.album.release_date),
                       color = factor(track.name),
                       fill = factor(track.artists),
                       alpha = track.popularity/100)) +
  
  geom_point() +
  
  labs(x = "Date Saved",
       y = "Release Date",
       title = "Discovery Scatterplot") +
  
  scale_alpha_identity() +
  
  myTheme 

ggplotly(msttsrdp)

```

```{r}

nowPlaying <- function() {
  
  np <- get_my_currently_playing()
  
  artists <- paste(np$item$artist$name, collapse = " & ")
  
  df <- data.frame("artist" = artists,
                   "song" = np$item$name,
                   "album" = np$item$album$name) |> 
    
  pivot_longer(cols = everything(),
               names_to = "Field",
               values_to = "Value")
  
  tf <- tempfile(fileext = ".jpg")
  
  download.file(get_my_currently_playing()$item$album$images$url[1],
                destfile = tf,
                mode = "wb",
                quiet = TRUE)
  
  ac <- image_read(tf)
  
  print(ac, 
        info = FALSE)
  
  if (df$Value[2] == df$Value[3]) {
    
    return(head(df, 2))
    
  } else {
    
    return(df)
    
  }
  
}

skip <- function() {
  
  np <- nowPlaying()
  
  print(paste("Skipping:",
              np$Value[2],
              "by",
              np$Value[1],
              sep = " "))
  
  print("Now playing: ")
  
  skip_my_playback()
  
  return(nowPlaying())
  
}

```

***

# Pull Audio Features for Saved Tracks

```{r, class.source="jhcv", message = FALSE}

# Get audio features for each track and combine into a single data frame
mstafdf <- imap_dfr(my_saved_tracks$track.id, ~{
  
  mstaf <- get_track_audio_features(.x)
  
  message(paste("Track ", .y, " out of ", length(my_saved_tracks$track.id), 
                " // ", round((.y/length(my_saved_tracks$track.id))*100, 2), "%"))
  
  mstaf
  
})

# Merge audio features with my_saved_tracks using left_join
my_saved_tracks <- left_join(my_saved_tracks, mstafdf, by = c("track.id" = "id"))

# View summary of my_saved_tracks
#kable(summary(my_saved_tracks))
```

# Music Map!

```{r, class.source="jhcv", message = FALSE}

mstafp <- ggplot(my_saved_tracks,
                 aes(x = valence,
                     y = energy,
                     color = factor(track.name),
                     fill = factor(track.artists),
                     alpha = track.popularity/100)) +
  
  geom_vline(xintercept = 0.5,
             color = "white") +
  geom_hline(yintercept = 0.5,
             color = "white") +
  
  geom_vline(aes(xintercept = mean(valence)),
             color = "cyan") +
  
  geom_hline(aes(yintercept = mean(energy)),
             color = "cyan") +
  
  geom_point() +
  
  labs(x = "Valence",
       y = "Energy",
       title = "Vibe Map") +
  
  scale_alpha_identity() +
  
  myTheme

ggplotly(mstafp)

```

# Create Averages of Audio Features for All Saved Tracks as a Time Series

```{r, class.source="jhcv", message = FALSE}

my_saved_tracks <- my_saved_tracks %>%
  arrange(added_at) %>%
  mutate(
    index = 1:n(),
    rmPop = cummean(track.popularity),
    rmValence = cummean(valence),
    rmEnergy = cummean(energy),
    rmTempo = cummean(tempo),
    rmDance = cummean(danceability),
    rmSpeech = cummean(speechiness),
    rmAcoustic = cummean(acousticness),
    rmInstrument = cummean(instrumentalness),
    rmLiveness = cummean(liveness)
  )

mstds <- my_saved_tracks |> mutate("ddate" = as.Date(added_at)) |>
  
  select(added_at,
         track.artists,
         track.name,
         track.album.name,
         key,
         mode,
         tempo,
         time_signature,
         track.popularity,
         energy,
         danceability,
         loudness,
         speechiness,
         acousticness,
         instrumentalness,
         liveness,
         rmPop,
         rmValence,
         rmEnergy,
         rmTempo,
         rmDance,
         rmSpeech,
         rmAcoustic,
         rmInstrument,
         rmLiveness)

```

```{r, class.source="jhcv", message = FALSE}

mstafa <- ggplot(my_saved_tracks,
                 aes(x = valence,
                     y = energy,
                     color = factor(track.name),
                     fill = factor(track.artists),
                     alpha = track.popularity/100)) +
  
  geom_vline(xintercept = 0.5,
             color = "white") +
  geom_hline(yintercept = 0.5,
             color = "white") +
  
  geom_vline(aes(xintercept = rmValence),
             color = "cyan") +
  
  geom_hline(aes(yintercept = rmEnergy),
             color = "cyan") +
  
  geom_point() +
  
  labs(x = "Valence",
       y = "Energy",
       subtitle = "Songs Saved by {frame_along}",
       title = "Music Taste Evolution!") +
  
  scale_alpha_identity() +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstafa, renderer = gifski_renderer())

```

```{r, class.source="jhcv"}

kdf <- data.frame("kn" = 0:11,
                  "kl" = c("C",
                           "C#",
                           "D",
                           "D#",
                           "E",
                           "F",
                           "F#",
                           "G",
                           "G#",
                           "A",
                           "A#",
                           "B"))

mdf <- data.frame("mn" = c(1, 0),
                  "ml" = c("major",
                           "minor"))

kmdf <- mstds |> select(key, mode) 

kmdf <- left_join(kmdf, kdf, by = c("key" = "kn"))

kmdf <- left_join(kmdf, mdf, by = c("mode" = "mn"))

kmdf$km <- ""

for (i in 1:nrow(kmdf)) {
  
  kmdf$km[i] <- paste(kmdf$kl[i], kmdf$ml[i], sep = " ")
  
}

kmdfs <- kmdf |> group_by(km) |> summarize("n" = n()) |> arrange(desc(n))

mkmp <- ggplot(kmdfs,
               aes(x = fct_reorder(km, n),
                   weight = n,
                   fill = fct_reorder(km, n))) +
  
  geom_bar() +
  
  geom_text(aes(y = n/2,
                label = km),
            color = "white") +
  
  labs(x = "Key & Mode",
       y = "Number of Tracks Saved",
       title = "My Favorite Keys & Modes") +
  
  myTheme +
  
  coord_flip() +
  
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())

ggplotly(mkmp)

```

```{r, class.source="jhcv", message = FALSE}

mstpa <- ggplot(mstds,
                aes(x = added_at,
                    y = rmPop)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Popularity",
       title = "Popularity Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstpa, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstpp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmPop)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Popularity",
       title = "Popularity Over Time") +
  
  myTheme

ggplotly(mstpp)

```

```{r, class.source="jhcv", message = FALSE}

mstva <- ggplot(mstds,
                aes(x = added_at,
                    y = rmValence)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Valence",
       title = "Valence Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstva, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstvp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmValence)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Valence",
       title = "Valence Over Time") +
  
  myTheme

ggplotly(mstvp)

```

```{r, class.source="jhcv", message = FALSE}

mstea <- ggplot(mstds,
                aes(x = added_at,
                    y = rmEnergy)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Energy",
       title = "Energy Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstea, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstep <- ggplot(mstds,
                aes(x = added_at,
                    y = rmEnergy)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Energy",
       title = "Energy Over Time") +
  
  myTheme

ggplotly(mstep)

```

```{r, class.source="jhcv", message = FALSE}

mstta <- ggplot(mstds,
                aes(x = added_at,
                    y = rmTempo)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Tempo",
       title = "Tempo Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstta, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

msttp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmTempo)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Tempo",
       title = "Tempo Over Time") +
  
  myTheme

ggplotly(msttp)

```

```{r, class.source="jhcv", message = FALSE}

mstda <- ggplot(mstds,
                aes(x = added_at,
                    y = rmDance)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Danceability",
       title = "Danceability Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstda, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstdp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmDance)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Danceability",
       title = "Danceability Over Time") +
  
  myTheme

ggplotly(mstdp)

```

```{r, class.source="jhcv", message = FALSE}

mstsa <- ggplot(mstds,
                aes(x = added_at,
                    y = rmSpeech)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Speechiness",
       title = "Speechiness Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstsa, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstsp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmSpeech)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Speechiness",
       title = "Speechiness Over Time") +
  
  myTheme

ggplotly(mstsp)

```

```{r, class.source="jhcv", message = FALSE}

mstaa <- ggplot(mstds,
                aes(x = added_at,
                    y = rmAcoustic)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Acousticness",
       title = "Acousticness Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstaa, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstap <- ggplot(mstds,
                aes(x = added_at,
                    y = rmAcoustic)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Acousticness",
       title = "Acousticness Over Time") +
  
  myTheme

ggplotly(mstap)

```

```{r, class.source="jhcv", message = FALSE}

mstia <- ggplot(mstds,
                aes(x = added_at,
                    y = rmInstrument)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Instrumentalness",
       title = "Instrumentalness Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstia, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstip <- ggplot(mstds,
                aes(x = added_at,
                    y = rmInstrument)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Instrumentalness",
       title = "Instrumentalness Over Time") +
  
  myTheme

ggplotly(mstip)

```

```{r, class.source="jhcv", message = FALSE}

mstla <- ggplot(mstds,
                aes(x = added_at,
                    y = rmLiveness)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Liveness",
       title = "Liveness Over Time") +
  
  myTheme +
  
  transition_reveal(added_at) +
  view_follow() +
  ease_aes("elastic-in-out")

animate(mstla, renderer = gifski_renderer())

```

```{r, class.source="jhcv", message = FALSE}

mstlp <- ggplot(mstds,
                aes(x = added_at,
                    y = rmLiveness)) +
  
  geom_line(color = "white") +
  
  labs(x = "Time",
       y = "Liveness",
       title = "Liveness Over Time") +
  
  myTheme

ggplotly(mstlp)

```

***

# Program Run Time

```{r, class.source="jhcv"}

timeFinished <- Sys.time()

timeElapsed <- timeFinished-timeStarted

tedf <- data.frame("Run Time" = paste(round(timeElapsed, 2),
                                      " minutes",
                                      sep = ""))

kable(tedf)

```

***